Preparing for Potential Closure of European Airspaces due to Re-entering Space Objects

### 2025-05-04 paper.qmd
editorial changes to effective expose area and rewritten impact probability

We assume the aircraft impact area to be much larger than the debris
object area so as to neglet the debris impact area.
We can also simplify the geometry of the aircraft and model it as
a rectangular solid with dimesions of wingspan ($W$), length ($L$) and
height ($H$). 
Now, if the aircraft were stationary, the area exposed to a debris impact would
be the top view, while if the debris were stationary it would be the front view.
In reality an aircraft en-route is moving (mainly) horizontally and the debris 
vertically [^balistic_descent], which which allows to express the effective
exposed area, $\varepsilon_k$, as

@@

where $s_{a}$ is the average aircraft speed, $s_{d_k}$ is the average fall speed
of the $k$-th debris object, $a_T$ area of the aircraft as viewed from the top,
$a_F$ area of the aircraft as viewed from the from the front of the aircraft.

In the calculations the following is assumed to be true ($t$ subscript for aircraft type)

@@

The effective exposed area now only depends on the aircraft type $t$ (dimensions
and cruise speed) and can be rewritten as:

@@

The values $W_t$, $H_t$ and $L_t$ have been retrieved from [@performance_database;@skybrary;@doc8643].

The impact probability for cell $h$ and single aircraft of type $t$ becomes:

>>>>

For the actual collision encounter between a re-entering object and an aircraft, we assume the aircraft impact area to be much larger than the debris object area.
Thus, the area spanned by the debris object is smaller than the aircraft dimensions. 
This assumption holds for practically all commercial aircraft within the scope of this study, and we can neglect the debris impact area.
Accordingly, we can also simplify the geometry of the aircraft and model it as a rectangular solid with dimesions of wingspan ($W$), length ($L$) and height ($H$).

Regarding the aircraft-debris encounter geometry, this yields to two conceptual considerations. 
<!-- TERMINOLOGY? rectangular solid -- impact --> top plane, front plane? Does the literture use "view"  -->
1.) if the aircraft were stationary, the area exposed to the debris impact would be the top view, and
2.) if the debris were stationary, it would be the front view.

Our focus is on commercial aircraft in flight, and - implicitly - operating at higher altitudes.
The en-route portion of a flight is characterised by the aircraft in straight and level flight, moving (mainly) horizontally.
At the same time, the movement of the re-entered debris object at regular en-route flight levels will be vertically [^balistic_descent].
This encounter geometry allows to express the effective exposed area, $\varepsilon_k$, as

@@

where 
* $s_{a}$ is the average aircraft speed, 
* $s_{d_k}$ is the average fall speed of the $k$-th debris object, 
* $a_T$ area of the aircraft as viewed from the top, and 
* $a_F$ area of the aircraft as viewed from the from the front of the aircraft.

This can be further instantiated for specific aircraft types.
In the calculations the following is assumed to be true ($t$ subscript for aircraft type)

@@

The effective exposed area now only depends on the aircraft type $t$ (i.e., its dimensions
and cruise speed) and can be rewritten as:

@@

These aircraft properties are well-known. 
While there does not exist a single openly available date base for these aircraft parameters, we collected the information from a series of sources and integrated them into a lookup dataset.
The values $W_t$, $H_t$ and $L_t$ have been retrieved from [@performance_database;@skybrary;@doc8643].

The impact probability for cell $h$ under consideration of a specific aircraft of type $t$ becomes:
